Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems

### 2022-10-12 marlin-paper.qmd
rough draft

---
title: "Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "`r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
bibliography: references.bib
params:
  run_name: ["v1.2"]
abstract: "Very rough cut. Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems. However, the design of spatial management strategies faces numerous challenges. Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways. We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research. We demonstrate the model's capabilities by focusing on three case studies on the conservation and food security impacts of marine protected areas (MPAs)."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---

@@

@@

Users can tune their fleets in one of two ways. First, they can specify a target exploitation rate *u* for each species in their model run. Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level. Second, they can specify a target depletion, defined as total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium depletion for each species matches the desired level.

@@

#### Dispersal

@@

@@

| State                  | Distribution          | Description                                     |
|------------------|------------------|-----------------------------------|
| $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$     |                       | Standard deviation around centroid              |
| Spatial catchability   |                       | Catchability varies in space                    |
| Spatial allocation     |                       | Fleet allocation strategy                       |
| F/M                    |                       | Fishing mortality relative to natural mortality |
| Adult diffusion rate   |                       |                                                 |
| Recruit diffusion rate |                       |                                                 |
| Hyperallometry         | $U(1,2)$              | Presence of hyperallometric fec                 |
| DD timing              |                       | Timing of density dependence                    |
| Seasonal Movement      | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

@@

```{r}
#| label: fig-coral-results
#| fig-cap: "Something something."
#| fig-height: 8
#| fig-width: 6

@@

* Growing need for more spatially explicit and ecosystem-based management and assessment

* `marlin` fills an important gap in the modeling literature
    * MPAs
    * Policy interaction
    * spatial-temporal MSE
    * Range shifts
    * make clear it's not just an MPA model

* Examples demonstrate this
    * total benefits and Distributional impacts in coral case
    * Something in the pelagic case
    * Complex interactions in the final case

* `marlin` and the methods within as a platform for future practical and academic bio-economic modeling needs

>>>>

---
title: "Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "`r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
bibliography: references.bib
params:
  run_name: ["v0.1"]
abstract: "Very rough cut. Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems. However, the design of spatial management strategies faces numerous challenges. Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways. We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research. We demonstrate the model's capabilities by focusing on three case studies on the conservation and food security impacts of marine protected areas (MPAs)."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

While users can adapt the general equations and ideas presented here as
they see fit, the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] and C++, integrated through the `Rcpp` package
@eddelbuettel2018 .

@@

Users can tune their fleets in one of two ways. First, they can specify
a target exploitation rate *u* for each species in their model run.
Taking all the other parameters of the model as given, `marlin` then
adjusts the catchability coefficients *q\_{f,s}* for each fleet *f* and
species *s* such that the equilibrium exploitation rate for each species
matches the desired level. Second, they can specify a target depletion,
defined as total spawning stock biomass under fishing divided by total
unfished spawning stock biomass, and the model will adjust the
catchability coefficients *q\_{f,s}* for each fleet *f* and species *s*
such that the equilibrium depletion for each species matches the desired
level.

@@

#### Movement

@@

One challenge is that the coefficients of the model matrices inside
@eq-movement can be hard to interpret. To facilitate this, `marlin`
allows users to specify a maximum number of cells that adults or
recruits diffuse in an year. The model will then tune the diffusion
parameter and habitat gradients to achieve this desired level of
movement.

@@

| State                  | Distribution          | Description                                     |
|-------------------|-------------------|-----------------------------------|
| $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$     |                       | Standard deviation around centroid              |
| Spatial catchability   |                       | Catchability varies in space                    |
| Spatial allocation     |                       | Fleet allocation strategy                       |
| F/M                    |                       | Fishing mortality relative to natural mortality |
| Adult diffusion rate   |                       |                                                 |
| Recruit diffusion rate |                       |                                                 |
| Hyperallometry         | $U(1,2)$              | Presence of hyperallometric fec                 |
| DD timing              |                       | Timing of density dependence                    |
| Seasonal Movement      | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

@@

```{r}
#| label: fig-coral-results
#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (top panel) for coral reef case study. Change in yield for each fleet as a function of MPA size and MPA design strategy (bottom panel) for coral reef case study."
#| fig-height: 8
#| fig-width: 6

@@

-   Growing need for more spatially explicit and ecosystem-based
    management and assessment

-   `marlin` fills an important gap in the modeling literature

    -   MPAs
    -   Policy interaction
    -   spatial-temporal MSE
    -   Range shifts
    -   make clear it's not just an MPA model
    -   Draw contrast to virtual species and STEPS

-   Examples demonstrate this

    -   total benefits and Distributional impacts in coral case
    -   Something in the pelagic case
    -   Complex interactions in the final case

-   `marlin` and the methods within as a platform for future practical
    and academic bio-economic modeling needs
